Skill estimation using TrueSkill Through Time: the \proglang{Julia}, \proglang{Python} and \proglang{R} packages

### 2021-04-23 article.tex
Paper TTT: revisi'on de la parte de ATP con Gustavo

\documentclass[article]{jss}
\usepackage[utf8]{inputenc}
% \usepackage{draftwatermark}
% \SetWatermarkText{Draft}
% \SetWatermarkScale{2}
\newif\ifen
\newif\ifes
\newcommand{\en}[1]{\ifen#1\fi}
\newcommand{\es}[1]{\ifes#1\fi}
\estrue
\usepackage[utf8]{inputenc}
\usepackage{paracol}
%\newcommand{\jan}{%
%  \en{January}%
%  \es{Enero}
%}

@@

@@

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{TrueSkill Through Time: the \proglang{Julia}, \proglang{Python} and \proglang{R} packages}
\Plaintitle{TrueSkill Through Time: the Julia, Python and R packages}
\Shorttitle{TrueSkill Through Time: the \proglang{Julia}, \proglang{Python} and \proglang{R} packages (Draft)}

@@

\en{In this last example we use the entire history of matches of the Association of Tennis Professionals (ATP).}
\es{En este último ejemplo utilizamos toda la historia de partidas de la Asociación de Tenistas Profesionales (ATP).}
%
\en{The database has \num{447000} games starting at year 1915 until 2020 with more than \num{19000} participating players and is publicly available.}
\es{La base de datos cuenta con \num{447000} mil partidas que van desde el año 1915 hasta el 2020, en las que participan \num{19000} jugadores.}
%
\en{The information is stored in a single \texttt{csv} file\footnote{File available at \nolinkurl{https://github.com/glandfried/tennis_atp/releases/download/atp/history.csv.zip}. An updated data source at \nolinkurl{https://github.com/JeffSackmann/tennis_atp}.} including both single and double matches: if the column \texttt{double} has the letter \texttt{t}, the game is a double match.}
\es{La información resumida en una archivo \texttt{csv}\footnote{Archivo disponible en \nolinkurl{https://github.com/glandfried/tennis_atp/releases/download/atp/history.csv.zip}. Una fuente de datos actualizada en \nolinkurl{https://github.com/JeffSackmann/tennis_atp}} contiene tanto partidas simples como dobles: si la columna \texttt{double} tiene la letra \texttt{t} la partida es doble.}
%
\en{Each game has an identifier (i.e. \texttt{match\_id}) and its tournament's round number (i.e. \texttt{round\_number}), where 0 represents the final game, 1 the semi-final, and so on.}
\es{Cada partida tiene un identificador (\texttt{match\_id}) y su número de ronda (\texttt{round\_number}), donde 0 representa la partida final, 1 la semi-final, etc.}
%
\en{The file also contains players' identifiers and names: for example column \texttt{w2\_id} is the identifier of the second player of the winning team and \texttt{l1\_name} is the name of the first player of the losing team.}
\es{The file also contains the identifiers and names of the players: por ejemplo \texttt{w2\_id} es el identificador del segundo jugador del equipo ganador y \texttt{l1\_name} es el nombre del primer jugador del equipo perdedor.}
%
\en{Finally, we have the name of the tournament (\texttt{tour\_name}), its identifier (\texttt{tour\_id}), the start date of the tournament (\texttt{time\_start}) and the type of surface (\texttt{ground}).}
\es{Finalmente tenemos el nombre del torneo (\texttt{tour\_name}), su identificado (\texttt{tour\_id}), la fecha de inicio del torneo (\texttt{time\_start}) y el tipo de piso (\texttt{ground}).}
%
\en{Here we only show the \proglang{Julia} version because it is more efficient than \proglang{Python} and \proglang{R} versions, which can be also found in the appendix \ref{sec:appendix_atp_code}.}
\es{Aquí mostramos solo la versión de \proglang{Julia} debido a que es la más eficiente, las versiones de \proglang{Python} y \proglang{R} se pueden entonrar en el anexo \ref{sec:appendix_atp_code}.}
%
\begin{lstlisting}[backgroundcolor=\color{julia},label=lst:atp, caption={\en{The history of the Association of Tennis Professionals}\es{La historia de la Asociación de Tenistas Profesionales}}, belowskip=0.1cm, aboveskip=0.1cm]
using CSV; using Dates
data = CSV.read("atp.csv")

times = Dates.value.(data[:,"time_start"] .- Date("1900-1-1")) 
composition = [ r.double == "t" ? [[r.w1_id,r.w2_id],[r.l1_id,r.l2_id]] : [[r.w1_id],[r.l1_id]] for r in eachrow(data) ]

h = History(composition=composition, times = times, sigma = 1.6, gamma = 0.036)
convergence(h,epsilon=0.01, iterations=10)

@@

\en{Temporal batches are defined as days elapsed since a reference date to the tournament start.}

@@

\en{When generating the list \texttt{composition} we discriminate whether the games are double or single based on the column \texttt{double}.}

@@

\en{The advantage of TrueSkill Through Time, over TrueSkill, is that by modeling the entire dynamic process it allows us to make comparisons of skills over time.}
\es{La ventaja de TrueSkill Through Time, respecto de TrueSkill, es que al modelar el proceso dinámico permite hacer comparaciones temporales de la habilidad.}

@@

\section*{Computational details}
% 
% \begin{leftbar}
% If necessary or useful, information about certain computational details
% such as version numbers, operating systems, or compilers could be included
% in an unnumbered section. Also, auxiliary packages (say, for visualizations,
% maps, tables, \dots) that are not cited in the main text can be credited here.
% \end{leftbar}

>>>>

\documentclass[article]{jss}
\usepackage[utf8]{inputenc}
% \usepackage{draftwatermark}
% \SetWatermarkText{Draft}
% \SetWatermarkScale{2}
\newif\ifen
\newif\ifes
\newcommand{\en}[1]{\ifen#1\fi}
\newcommand{\es}[1]{\ifes#1\fi}
\entrue
\usepackage[utf8]{inputenc}
\usepackage{paracol}
%\newcommand{\jan}{%
%  \en{January}%
%  \es{Enero}
%}

@@

\hypersetup{
    colorlinks,
    linkcolor={black!50!black},
    citecolor={black!50!black},
    urlcolor={black!80!black}
}

@@

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{Skill estimation using TrueSkill Through Time: the \proglang{Julia}, \proglang{Python} and \proglang{R} packages}
\Plaintitle{TrueSkill Through Time: the Julia, Python and R packages}
\Shorttitle{TrueSkill Through Time: the \proglang{Julia}, \proglang{Python} and \proglang{R} packages (Draft)}

@@

\en{In this last example, we analyze the complete history of the Association of Tennis Professionals (ATP) registered matches.}
\es{En este último ejemplo utilizamos toda la historia de partidas de la Asociación de Tenistas Profesionales (ATP).}
%
\en{The database has \num{447000} games starting at year 1915 until 2020 with more than \num{19000} participating players and is publicly available.}
\es{La base de datos cuenta con \num{447000} mil partidas que van desde el año 1915 hasta el 2020, en las que participan \num{19000} jugadores.}
%
\en{The information is obtained and stored in a single \texttt{csv} file\footnote{Available at \url{https://github.com/glandfried/tennis_atp/releases/download/atp/history.csv.zip}. Additional data at Jeff Sackmann's site: \url{https://github.com/JeffSackmann/tennis_atp}.} including both single and double matches: if the column \texttt{double} has the letter \texttt{t}, the game is a double match.}
\es{La información resumida en una archivo \texttt{csv}\footnote{Archivo disponible en \nolinkurl{https://github.com/glandfried/tennis_atp/releases/download/atp/history.csv.zip}. Una fuente de datos actualizada en \nolinkurl{https://github.com/JeffSackmann/tennis_atp}} contiene tanto partidas simples como dobles: si la columna \texttt{double} tiene la letra \texttt{t} la partida es doble.}
%
\en{Each game has an identifier (i.e. \texttt{match\_id}) and its tournament's round number (i.e. \texttt{round\_number}), where 0 represents the final game, 1 the semi-final, and so on.}
\es{Cada partida tiene un identificador (\texttt{match\_id}) y su número de ronda (\texttt{round\_number}), donde 0 representa la partida final, 1 la semi-final, etc.}
%
\en{The file also contains players' identifiers and names: for example column \texttt{w2\_id} is the identifier of the second player of the winning team and \texttt{l1\_name} is the name of the first player of the losing team.}
\es{The file also contains the identifiers and names of the players: por ejemplo \texttt{w2\_id} es el identificador del segundo jugador del equipo ganador y \texttt{l1\_name} es el nombre del primer jugador del equipo perdedor.}
%
\en{Finally, we have the name of the tournament (\texttt{tour\_name}), its identifier (\texttt{tour\_id}), the start date of the tournament (\texttt{time\_start}) and the type of surface (\texttt{ground}).}
\es{Finalmente tenemos el nombre del torneo (\texttt{tour\_name}), su identificado (\texttt{tour\_id}), la fecha de inicio del torneo (\texttt{time\_start}) y el tipo de piso (\texttt{ground}).}
%
\en{Here we only show the \proglang{Julia} code because it is far more efficient than \proglang{Python} and \proglang{R} versions (in section~\ref{sec:computationDetails} we show performance comparison of the three versions) which can also be found in the appendix~\ref{sec:appendix_atp_code}.}
\es{Aquí mostramos solo la versión de \proglang{Julia} debido a que es la más eficiente, las versiones de \proglang{Python} y \proglang{R} se pueden entonrar en el anexo \ref{sec:appendix_atp_code}.}
%
\begin{lstlisting}[backgroundcolor=\color{julia},label=lst:atp, caption={\en{The history of the Association of Tennis Professionals}\es{La historia de la Asociación de Tenistas Profesionales}}, belowskip=0.1cm, aboveskip=0.1cm]
using CSV; using Dates
data = CSV.read("atp.csv")

dates = Dates.value.(data[:,"time_start"] .- Date("1900-1-1")) 
matches = [ r.double == "t" ? [[r.w1_id,r.w2_id],[r.l1_id,r.l2_id]] : [[r.w1_id],[r.l1_id]] for r in eachrow(data) ]

h = History(composition = matches, times = dates, sigma = 1.6, gamma = 0.036)
convergence(h, epsilon = 0.01, iterations = 10)

@@

\en{We assume the skill of each player does not change during a tournament and we add uncertainty (based on $\gamma$ parameter) before each tournament according to the elapsed time between competitions.}

@@

\en{When generating the list \texttt{composition} we discriminate whether the games are doubles or singles using the column \texttt{double}.}

@@

% \en{The advantage of TrueSkill Through Time, over TrueSkill, is that by modeling the entire dynamic process it allows us to make comparisons of skills over time.}
% \es{La ventaja de TrueSkill Through Time, respecto de TrueSkill, es que al modelar el proceso dinámico permite hacer comparaciones temporales de la habilidad.}

@@

\section*{Computational details}\label{sec:computationDetails}
% 
% \begin{leftbar}
% If necessary or useful, information about certain computational details
% such as version numbers, operating systems, or compilers could be included
% in an unnumbered section. Also, auxiliary packages (say, for visualizations,
% maps, tables, \dots) that are not cited in the main text can be credited here.
% \end{leftbar}
